\documentclass[12pt,letterpaper]{report}
\usepackage[margin=1in,includeheadfoot]{geometry}
\usepackage[pdftex,
pdfauthor={Cheuk Chuen Siow},
pdftitle={Concern-Oriented Use Case Maps},
pdfsubject={Computer Science},
pdfkeywords={CORE, UCM},
hidelinks,bookmarksnumbered]{hyperref}
\usepackage[figure,table]{hypcap}
\usepackage[nottoc]{tocbibind}
\usepackage[titletoc]{appendix}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{todonotes}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{subfig}
\usepackage{enumitem}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\rhead{\textit{\leftmark}}
\cfoot{\thepage}
\renewcommand{\headrulewidth}{0pt}

\usepackage[backend=biber,sorting=none]{biblatex}
\addbibresource{references.bib}

\usepackage{graphicx}
\graphicspath{ {images/} }

\usepackage{setspace}
\doublespacing

\usepackage{datetime}
\newdateformat{monthyeardate}{\monthname[\THEMONTH] \THEYEAR}
\newdateformat{yeardate}{\THEYEAR}

\usepackage{sectsty}
\chapterfont{\vspace*{-17.5mm}\large\sc\centering}
\chaptertitlefont{\centering}
\subsubsectionfont{\centering}

\usepackage{chngcntr}
\counterwithout{footnote}{chapter}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\newcommand*{\cls}{\fontfamily{cmss}\selectfont}

\newcommand{\jkReplace}[2]{\textcolor{red}{JK replaced: }\textcolor{blue}{#1}{\textcolor{red}{\ by }\textcolor{green}{#2}}}

%----------------
% Title Page
%----------------
\begin{document}
	
	\begin{titlepage}
		\begin{center}
			\vspace*{1in}
			
			\Huge
			Concern-Oriented Use Case Maps
			
			\vfill
			
			\large
			Cheuk Chuen Siow\\
			School of Computer Science\\
			McGill University, Montréal
			
			\vfill
			
			\monthyeardate\today
			
			\vfill
			
			A thesis submitted to McGill University in partial fulfillment of the requirements of the degree of Master of Science
			
			\vfill
			
			\textcopyright{} Cheuk Chuen Siow, \yeardate\today
			
			\vspace*{1in}
		\end{center}
	\end{titlepage}
	
	%----------------
	% Front Matter
	%----------------
	\clearpage
	\pagenumbering{roman}
	\setcounter{page}{2}
	
	\chapter*{Abstract}
	\markboth{\uppercase{Abstract}}{}
	\addcontentsline{toc}{chapter}{Abstract}
	
	Concern-Oriented Reuse (CORE) is a reuse paradigm that extends model-driven engineering with advanced modularization, goal modeling, and software product lines. Previous work enables modeling with CORE at the design level using Reusable Aspect Models (RAM). Requirements elicitation is also a crucial aspect of software development process, and one of the visual notation that expresses use cases as graphical workflows is Use Case Maps (UCM). UCM bridges the gap between requirements and design, and is part of the User Requirements Notation (URN) for scenario modeling. This thesis addresses the need for enabling scenario modeling in CORE. Based on Aspect-Oriented Use Case Maps (AoUCM), we introduce a novel technique that applies advanced separation of concerns for model-driven requirements elicitation with use cases---Concern-Oriented Use Case Maps (CoUCM). We define a metamodel for CoUCM that derives from the CORE metamodel, and formulate the weaving algorithm for CoUCM model composition. We then implement CoUCM in the TouchCORE tool as proof of concept. We present a working application of scenario modeling with TouchCORE, in which we further validate our implementation through case studies and workflow patterns. Validation shows that CoUCM is able to model certain requirements concerns in a reusable way, and that they can then easily be applied to different reuse contexts.
	
	\clearpage
	
	\chapter*{Abrégé}
	\markboth{\uppercase{Abrégé}}{}
	\addcontentsline{toc}{chapter}{Abrégé}
	
	\emph{Concern-Oriented Reuse} (CORE)\footnote{Traduit en Fran{\c c}ais: "La R{\'e}utilisation Orient{\'e}-Pr{\'e}occupations"} est une nouvelle approche de r{\'e}utilisation qui s'appuie sur l'ing{\'e}nierie dirig{\'e}e par les mod{\`e}les (IDM), la modularisation avanc{\'e}e, la mod{\'e}lisation des besoins, et les lignes de produits logiciels. Jusqu'{\`a} pr{\'e}sent, CORE permettait seulement la mod{\'e}lisation de conceptions avec le langage de mod{\'e}lisation \emph{Reusable Aspect Models} (RAM). Cependant, la mod{\'e}lisation des besoins est une partie fondamentale de l'approche IDM. \emph{Use Case Maps} (UCM) est un langage de mod{\'e}lisation qui permet d'exprimer les cas d'utilisations visuellement en tant que flux de travails. UCM permets de formaliser les cas d'utilisations textuels et donc informels, et ainsi passer plus facilement des mod{\`e}les de besoins au mod{\`e}les de conc{\'e}ption. Ce travail de th{\`e}se int{\`e}gre UCM avec l'approche de r{\'e}utilisation CORE. En s'inspirant de la t{\'e}chnique Aspect-Oriented Use Case Maps (AoUCM), nous proposons une extension de UCM orient{\'e}-pr{\'e}occupations nomm{\'e}e \emph{Concern-Oriented Use Case Maps} (CoUCM). Nous d{\'e}finissons un meta mod{\`e}le pour CoUCM qui s'int{\`e}gre avec le m{\'e}ta mod{\`e}le de CORE, et nous proposons des algorithmes de tissage qui permettent la composition de mod{\`e}les UCM venant de diff{\'e}rents pr{\'e}occupations ou charact{\'e}ristiques. Une impl{\'e}mentation prototype de CoUCM dans l'outil de mod{\'e}lisation TouchCORE est {\'e}galement d{\'e}crite. Pour valider l'expressivit{\'e} de notre solution, nous montrons comment CoUCM s'applique {\`a} plusieurs cas d'utilisation, qui d{\'e}montrent que CoUCM permet de modulariser et rendre r{\'e}utilisable certaines pr{\'e}occupations, tel que par exemple les diff{\'e}rents int{\'e}ractions de payment.	
	
	\clearpage
	
	\chapter*{Acknowledgements}
	\markboth{\uppercase{Acknowledgements}}{}
	\addcontentsline{toc}{chapter}{Acknowledgements}
	
	Foremost, I would like to express my sincere gratitude to Professor Jörg Kienzle and Professor Gunter Mussbacher, my research supervisors, for their patient guidance and insightful comments of this research work. This thesis would not be possible without their supervision.
	
	I thank my fellow colleagues from the Software Engineering Lab for their encouragement and stimulating discussions that challenged me intellectually.
	
	Last but not least, I would like to thank my family and friends around the world for their support throughout my study.
	
	% Content
	\renewcommand{\contentsname}{Table of Contents}
	\tableofcontents
	\listoffigures
	%\listoftables
	\listofalgorithms
	\addcontentsline{toc}{chapter}{List of Algorithms}
	
	\clearpage
	
	\chapter*{List of Acronyms}
	\markboth{\uppercase{List of Acronyms}}{}
	\addcontentsline{toc}{chapter}{List of Acronyms}
	
	\begin{description}[noitemsep,style=multiline,leftmargin=3cm]
		\item[ADORE] Activity moDel supOrting oRchestration Evolution
		\item[AoUCM] Aspect-Oriented Use Case Maps
		\item[AoURN] Aspect-Oriented User Requirements Notation
		\item[BPEL] Business Process Execution Language
		\item[BPMN] Business Process Model and Notation
		\item[CORE] Concern-Oriented Reuse
		\item[CoUCM] Concern-Oriented Use Case Maps
		\item[EMF] Eclipse Modeling Framework
		\item[GRL] Goal-oriented Requirements Language
		\item[GUI] Graphical User Interface
		\item[ITU] International Telecommunication Union
		\item[MDE] Model-Driven Engineering
		\item[MVC] Model-View-Controller
		\item[RAM] Reusable Aspect Models
		\item[SOA] Service-Oriented Architecture
		\item[SoC] Separation of Concerns
		\item[SPL] Software Product Lines
		\item[UCM] Use Case Maps
		\item[URN] User Requirements Notation
		\item[XML] Extensible Markup Language
	\end{description}
	
	\clearpage
	\pagenumbering{arabic}
	
	%----------------
	% Main Body
	%----------------
	\chapter{Introduction}
	\input{chapters/1-introduction}
	
	\chapter{Background} \label{ch:2}
	\input{chapters/2-background}
	
	\chapter{Adding Support for UCM to CORE} \label{ch:3}
	\input{chapters/3-support}
	
	\chapter{Validation} \label{ch:4}
	\input{chapters/4-validation}
	
	\chapter{Conclusion} \label{ch:5}
	\input{chapters/5-conclusion}
	
	%----------------
	% End Matter
	%----------------
	\renewcommand\bibname{References}
	\printbibliography[heading=bibintoc]
	
	\begin{appendices}
		
		\chapter{Complete Metamodels} \label{ch:A}
		\input{chapters/A-appendix}
		
	\end{appendices}
	
\end{document}
